---
title: "Notes for Assignment 1"
author: "Haviland Wright"
date: "September 15, 2018"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo = TRUE)
```


## 1 Calculate $p_k$

Given the stated mean value of 2 misprints per page, assume the distribution of $M$, the number of misprints on a page, is $Poisson(\lambda = 2)$ 

Conditioned on $k$, a limit to be determined, the probability $p_k$ of more than $k$ misprints on a page is:  

$p_k = P(M > k) = 1 - P(M \leq k) =$ 1 - ppois($k$, lambda = 2).  


## 2 Calculate $P(T \leq n | k)$

Probability of $T$ no more than $n$ pages with more than $k$ misprints is:  

$P(T \leq n) =$ pbinom($n$, 50, $p_k$).  

\vspace{.2in}

## 3 Visualize $P(n \leq k)$

For a given $k$, the probability mass and cumulateive distribution functions of errors per page can be visualized as follows:
\vspace{.2in}
```{r }

options(digits = 3)
options(scipen = 999)

## produce a dataframe containing probabilities
prob <- as.data.frame(dpois(x = 0:13, lambda = 2))
prob <- cbind(prob, cumsum(prob)) ## these probabilities = ppois(0:10,lambda=2)
prob <- cbind(prob, 1 - prob[,2])

colnames(prob) <- c("prob", "cprob", "tail_prob")

## barplot with mass function and CDF
df.bar <- barplot(prob$prob,  
                  main="Poisson mass function, lambda = 2",
                  names.arg = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13"),
                  xlab = "X: Number of misprints on a page", 
                  ylab = "Probabilty: P(x)",ylim = c(0,1.1), col="red")

lines(x=df.bar, y=prob$cprob, col="black")
points(x=df.bar, y=prob$cprob, pch=16, col="blue")
abline(h=1, lty=2, col = "cyan")

```
\vspace{.3in}  

## 4 Key Values Table

A table of the first 11 probabilities provides the tail area 
probabilities, $p_k = P(M > k)$.
\vspace{.2in}


```{r}

k <- 0:13

prob <- cbind(k, prob)

## format data frame for output as a table
p.tab <- kable(prob)
kable_styling(p.tab, bootstrap_options = "striped", full_width = FALSE, position = "left")

```
\newpage

## 5  Visualize  $P(T = n | k = 1, 3, 5)$

In an n rows by k columns table, each column will reflect different parameters

```{r}

# Setup for multiple plots
par(mfrow=c(3,1))

# calculate probability functions for display, make barplots.
f1 <- dbinom(x = 0:50, size = 50, prob = .594)
b1 <- barplot(f1, main="k = 1", axes = TRUE)


f3 <- dbinom(x = 0:50, size = 50, prob = .143)
barplot(f3, main="k = 3")

f5 <- dbinom(x = 0:50, size = 50, prob = .017)
barplot(f5, main="k = 5")


```


## 6  Example: Column from n x k table

```{r}
# explanition for the model 

#when n pages are selected from 50 pages book, it follows binomial distribution. 
#for calculation of possibility for one pages that has k misprints, 
#it follows the poisson distribution.

table <- as.data.frame(matrix(ncol = 5,nrow=50, dimnames = list(c(1:50),c(1:5))))
for ( i in 1:50) {  #loop 1 for the whole 50 pages variable 
  for (j in 1:5){ #loop 2 for the k = 1,2,3,4,5 
    
    p <- ppois(j,lambda = 2,lower.tail = FALSE) 
    table[i,j] = pbinom(i, size=50, prob=p) 
  }
}

# to adjust the table contents 
kableExtra::kable(digits = 3,x =table,row.names = TRUE)


```



